---TEST
title: "Case Study 2 - Group 4"
author:
- Annika Janson h11829506
- Jan Beck h11814291
- Franz Uchatzi
date: "8.11.2020"
output:
  pdf_document: default
  html_document:
    df.print: paged
header-includes: \renewcommand{\and}{\\}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
chick <- read.csv("chicken.csv")
```

## 3 Data Analysis

### 3.1

```{r, echo=FALSE}

price <- chick$pchick
demand <- chick$consum

plot(x = price,
     y = demand,
     xlab="Price", #Beschriftung x-Achse
     ylab="Demand", #Beschriftung y-Achase
     main="Scatterplot") #Beschriftung Grafik
     #abline(lm(y~x), col="red"), #regression line (y~x)
     #lines(lowess(x,y), col="blue")) # lowess line (x,y)
      

```
- positive, but economics theory would suggest a negative correlation


### 3.2 

```{r, echo=FALSE}
reg1 <- lm(log(consum) ~ log(income) + log(pchick) + log(pbeef) + log(ppork), data=chick) # attention logs!
summary(reg1)

m1_B0 <- as.vector(coef(reg1)[1]) # intercept
m1_B1 <- as.vector(coef(reg1)[2]) # beta1hat income
m1_B2 <- as.vector(coef(reg1)[3]) # beta2hat chicken
m1_B3 <- as.vector(coef(reg1)[4]) # beta3hat beef
m1_B4 <- as.vector(coef(reg1)[5]) # beta4hat pork

```


- result suggest that the higher the price for chicken the lower demand for it (negative correlation)
- compatible with economic theory
- extra question: 1) so that we can interpret it as elasticities (% changes) 2) to make the correlation "more linear"


### 3.3

- A 1% increase of income leads to a __`r m1_B1`__% change in demand.


### 3.4

- A 2% increase of income leads to a __`r 2*m1_B1`__% change in demand.


### 3.5

```{r, echo=FALSE}

log_demand <- predict(reg1, data.frame(income=2200, pchick=50, ppork=170, pbeef=312)) 


```

- The log demand is __`r log_demand`__ 

- The demand is __`r exp(log_demand)`__ 

### 3.6

```{r, echo=FALSE}

fitted <- fitted(reg1)
r <- cov(fitted, demand) / (sd(fitted) * sd(demand))

SStot <- sum( (demand - mean(demand)) ^ 2 )
SSres <- sum( ( demand - fitted )^2)

r2 <- 1 - SSres/SStot

# we can use the regression summary output to check if this is correct
```

### 3.7

### 3.8


\newpage
## 4 Theory

### 4.1 

$$
\begin{aligned}
A^2 + B^2 &= C^2 \\
A^2 &= C^2 - B^2
\end{aligned}
$$


### 4.2