---
title: "Case Study 3 - Group 4"
author:
- Annika Janson h11829506
- Jan Beck h11814291
- Franz Uchatzi
date: "29.11.2020"
output:
  pdf_document: default
  html_document:
    df.print: paged
header-includes: 
 - \usepackage{dcolumn}
 - \renewcommand{\and}{\\}
---


```{r setup, include=FALSE}
library(stargazer)
library(extrafont)
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
wages <- read.csv("WD1.csv")
N <- nrow(wages)
```

## 2 Descriptive statistics

```{r, results='asis', echo=FALSE}
stargazer(wages, omit = c("X"), header=FALSE, type='latex', summary.stat=c('mean', 'sd', 'median', 'min', 'max'), title='Summary statistics', align=TRUE)
```
\newpage
### 2.1

The average wage is USD __`r round(mean(wages$wage), 2)`__ and the median wage is USD __`r median(wages$wage)`__.

```{r, echo=FALSE}
hist(wages$wage, main="Histogram of wage", xlab="Wage in USD")
legend("topright", legend=c("Mean","Median"), lwd=c(2,2), col=c("red","blue"))
abline(v = mean(wages$wage), col = "red", lwd = 2)
abline(v = median(wages$wage), col = "blue", lwd = 2)

```

In the histogram we see that the distribution is right-skewed with a few observations exceeding USD 3000. Most observed values are concentrated around an interval of USD ±500 above and below the mean. Median and mean are fairly close to each other with the median being slightly higher due to the large outliers.

### 2.2

The proportion of workers working more than 40 hours a week is `r length(which(wages$hours > 40)) / N * 100`%.

### 2.3

```{r, echo=FALSE}
#table(wages$educ),decreasing=TRUE)
```

The most common number of years of education among the workers is `r as.vector(sort(table(wages$educ),decreasing=TRUE)[1])[1]`.

### 2.4

```{r, echo=FALSE}
#sort(table(wages$sibs),decreasing=TRUE)
```

No, the most frequent sibling pattern is having `r names(sort(table(wages$sibs),decreasing=TRUE))[1]` sibling.

\newpage
## 3 Data modelling

### 3.1

### 3.2

### 3.3

### 3.4

### 3.5.1

### 3.5.2

### 3.5.3

### 3.5.4

### 3.5.5

### 3.5.6



\newpage
## 4 Simulation Study

### 4.1 

### 4.2

### 4.3