---
title: "Case Study 1 - Group 4"
author:
- Annika Janson h11829506
- Jan Beck h11814291
- Franz Uchatzi
date: "18.10.2020"
output: 
  pdf_document: default
  html_document:
    df.print: paged

header-includes:
  - \renewcommand{\and}{\\}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
engelsheet <- read.csv("engel.csv")
```

## 1 Data Analysis

### 1.1 

```{r, echo=FALSE}

income <- engelsheet$income
food_exp <- engelsheet$foodexp

plot(x = income,
     y = food_exp,
     xlab="Houshold income", #Beschriftung x-Achse
     ylab="Houshold food expenditure", #Beschriftung y-Achase
     main="Scatterplot") #Beschriftung Grafik
     #abline(lm(y~x), col="red"), #regression line (y~x)
     #lines(lowess(x,y), col="blue")) # lowess line (x,y)
      



```

In this scatter plot on the x-axis we see the income per houshold. On the y-axis we see the foodexpenditure per houshold. The unit for both variables is 1 franc per value. We can see that there is a positive correlation between houlshold income and food expenditure. The more income a houshold has the more is spent on food in general. However, we can observe that there exists a certain threshold for food expenditure even with the highest income at about 2000 franc. Additionally, there are some outliers with an income of 4957.8 and food expenditure of only 1827.2. Another outlier has an income of approximately 2600 and food expenditure of 900. We generally have less observations for higher income households than lower ones.


### 1.2 

```{r pressure, echo=FALSE}
log_food_exp <- log(food_exp)
log_income <- log(income)

plot(x = log_income,
     y = log_food_exp,
     main ="log-log plot",
     )

```

The graphed correlation looks a lot more linear compared to the previous non-logged one. The outliers are less extreme and most observations crowd around the center of the plot rather than the bottom.


### 1.3

```{r, echo=FALSE}

hist(income,
     xlab="Income per houshold",
     ylab="Freq",
     breaks = 50)

hist(food_exp,
     main="Histogram of food expenditure",
     xlab="Food expenditure per houshold",
     ylab="Freq",
     breaks= 50)

summary(engelsheet)


```

The last output is the summary of the variables income and foodexp. For each variable we can see different key information such as Min and Max values. 


### 1.4 Linear regression Model 1

```{r, echo=FALSE}
reg1 <- lm(food_exp ~ income) #lineare regression mit standardvariablen
m1_B0 <- as.vector(coef(reg1)[1]) # intercept
m1_B1 <- as.vector(coef(reg1)[2]) # beta1hat

summary(reg1)

```

The summary of the linear regression of Model 1, where food expenditure is the dependent variable and household income is the explanatory variable, shows us that the estimate for the intercept $\hat{\beta}_{0}$ is __`r m1_B0`__ and for the slope $\hat{\beta}_{1}$ it is __`r m1_B1`__. Because $\hat{\beta}_{1}$ > 0, changes of X and Y move into the same direction. 


### 1.5

```{r, echo=FALSE}

#The y-intercept of a line is the value of y where the line crosses the y-axis. In other words, it is the value of y when the value of x is equal to 0. Sometimes this has true meaning for the model that the line provides, but other times it is meaningless.


```

According to the estimates of Model 2, food expenditure is expected to increase by __`r round(m1_B1, 2)`__ franc for every additional franc of income.


### 1.6 Linear regression Model 2

```{r, echo=FALSE}

reg2 <- lm(log_food_exp ~ log_income) #lineare regression mit "log variablen"

m2_B0 <- as.vector(coef(reg2)[1]) # intercept
m2_B1 <- as.vector(coef(reg2)[2]) # beta1hat

summary(reg2)


```

The summary of the linear regression of Model 2, where logged food expenditure is the dependent variable and logged household income is the explanatory variable, shows us that the estimated is intercept $\hat{\beta}_{0}$ is __`r m2_B0`__ and  the slope $\hat{\beta}_{1}$ is __`r m2_B1`__.


### 1.7

According to Model 2, we can say that for every percentage increase of income, food expenditure is expected to increase by __`r round(m2_B1, 2)`%__.


### 1.8

```{r, echo=FALSE}

pred_foodexp1 <- m1_B0 + m1_B1 * 10000
pred_foodexp2 <-exp( m2_B0 + m2_B1 * log(10000) ) # convert back from logged values

```

According to Model 1 the expected food expenditure of a household with an income of 10,000 franc is __`r round(pred_foodexp1, 2)`__ franc.

According to Model 2 the expected food expenditure of a household with an income of 10,000 franc is __`r round(pred_foodexp2, 2)`__ franc.


## 2 Theory

### 2.1 

```{r}
reg1_2  <- lm(income ~ food_exp) # für 2.1 muss man die dependent varaible Y mit der explanatory variable X vertauschen 

summary(reg1_2)

cov1 <- cov(income, food_exp)  # hab ich unnötig berechnet wie ich später gesehen habe,
# da Beta1 vom vorherigen Beispiel eh schon gegeben war
var1 <- var(income)
beta1 <- (cov1/var1)

print(beta1)
print(1/beta1)
```

In this summary we set the dependent variable to X and the explanatory variable to Y. We can see that $\frac{1}{0.4851784} = 2.061097$ and therefore the argument $\gamma_1 = \frac{1}{\beta_1}$ is false. $\gamma_1 = 1.71146$. 


--------- +++ Bin mir nicht sicher ob ich die Angabe von dem Aufgabe verstehe, in Bezug auf den "Hint".. für was soll das gut sein? Soll man sich da nochmal Gamma1 ausrechnen..

<!-- We have to do another regression -->






### 2.2


$I \rightarrow II$


$$y = \beta_0+\beta_1X+u\,\,\,\,\,\,\,\,\,\,\,\,\ \mathbb{E}[u|X]=0\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,$$
$$\rightarrow u=y-\beta_0-\beta_1X\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,$$
$$u \rightarrow \mathbb{E}[u|X]=\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,$$


$$=\mathbb{E}[y-\beta_0-\beta1X|X]=\mathbb{E}[Y|X]-\mathbb{E}[\beta_0|X]- \mathbb{E}[\beta_1X|X] $$
$$0=\mathbb{E}[Y|X]- \beta_0\mathbb{E}[1|X]- \beta_1\mathbb{E}[X|X]\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,$$



$$0=\mathbb{E}[Y|X]- \beta_0-\beta_1X\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,$$
$$\mathbb{E}[Y|X] = \beta_0 + \beta_1X\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,$$



$II \rightarrow I$


$$\mathbb{E}[Y|X]=\beta_0 + \beta_1X\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,$$
$$\mathbb{E}[\beta_0 + \beta_1X+u|X]=\beta_0 + \beta_1X\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,$$
$$\mathbb{E}[\beta_0|X]+\mathbb{E}[\beta_1X|X]+\mathbb{E}[u|X]=\beta_0 + \beta_1X\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,$$
$$\beta_0 + \beta_1X+\mathbb{E}[u|X]=\beta_0 + \beta_1X\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,$$
$$\rightarrow \mathbb{E}[u|X]=0\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,\,$$
